The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-23 script/de/chapters/ellipticcurves.tex
Script[de+en]: Small additions like research about Benford's law and primes.

Sage-Notebooks werden im Browser nach einem Logon-Vorgang aufgerufen%
\footnote{%
Hat man Sage auf einem eigenen (Unix-)Server installiert, muss man auf der Sage-Kommandozeile 
erst den Befehl \verb#notebook()# aufrufen.
},\footnote{%
Das ECC-Notebook von Frau Massierer benötigt die KASH3-Bibliothek: Deshalb muss (z.B. für Sage 4.1)
das Package \glqq kash3-2008-07-31.spkg\grqq~compiliert worden sein.
}.

Das ECC-Notebook von Massierer\footnote{%
Loggt man sich auf einem fremden, öffentlichen Server wie
\url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/} oder
\url{http://www.sagenb.org/} ein, und liegen dort die Worksheets als
\glqq Published Worksheets\grqq~schon vor, kann man folgendermaßen vorgehen:\\
- Starten des Sage-Notebooks in einem Browser.

@@

},\footnote{%
  Seit 2008 steht dieses Tutorial auf
    \url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}.\\
  Um sich in Siegen einzuloggen, muss man den Port 8000 und Cookies erlauben.
  %  -> Zum Ansehen: (siehe Massierers Mail vom 05.09.2008 12:52)
}
besteht aus 8 Kapiteln (\glqq Titelseite\grqq~mit Inhaltsverzeichnis plus 7 Kapitel)
und zielt darauf ab, dass selbst ein Einsteiger versteht, was Elliptische Kurven
sind:

>>>>

Sage-Notebooks werden im Browser nach einem Logon-Vorgang aufgerufen%
\footnote{%
Hat man Sage auf einem eigenen (Unix-)Server installiert, muss man auf der Sage-Kommandozeile 
erst den Befehl \verb#notebook()# aufrufen.
}${}^,$\footnote{%
Das ECC-Notebook von Maike Massierer benötigt die KASH3-Bibliothek: Deshalb muss (z.B. für Sage 4.2.1)
das Package \glqq kash3-2008-07-31.spkg\grqq~installiert worden sein (Befehl \verb#sage -i#).
}.

Das ECC-Notebook\index{Elliptische Kurven!ECC-Notebook}\index{Massierer, Maike} von
Massierer\footnote{%
Loggt man sich auf einem fremden, öffentlichen Server wie
\url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/} oder
\url{http://www.sagenb.org/} ein, und liegen dort die Worksheets als
\glqq Published Worksheets\grqq~schon vor, kann man folgendermaßen vorgehen:\\
- Starten des Sage-Notebooks in einem Browser.

@@

}${}^,$\footnote{%
  Seit 2008 steht dieses ECC-Notebook auf
    \url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}.\\
  Um sich in Siegen einzuloggen, muss man den Port 8000 und Cookies erlauben.
  %  -> Zum Ansehen: (siehe Massierers Mail vom 05.09.2008 12:52)
}
besteht aus 8 Teilen (\glqq Titelseite\grqq~mit Inhaltsverzeichnis plus 7 Kapitel)
und zielt darauf ab, dass selbst ein Einsteiger versteht, was Elliptische Kurven
sind (es ist nur auf Englisch vorhanden):

### 2009-11-23 script/de/chapters/primes.tex
Script[de+en]: Small additions like research about Benford's law and primes.

\item Bernhard Riemann\footnote{%
        Bernhard Riemann\index{Riemann, Bernhard}, deutscher Mathematiker,
        17.9.1826$-$20.7.1866. }
    stellte eine bisher unbewiesene, aber auch nicht widerlegte Formel
    für die Verteilung von Primzahlen auf, die die Abschätzung 
    weiter verbessern würde.
\end{itemize}

@@

Primzahlzwillinge sind Primzahlen, die den Abstand 2 voneinander haben, 
zum Beispiel 5 und 7, oder 101 und 103,
oder $1.693.965 \cdot 2^{66.443} \pm 1.$.

>>>>

\item Bernhard Riemann\footnote{%
        Bernhard Riemann\index{Riemann, Bernhard}, deutscher Mathematiker,
        17.9.1826$-$20.7.1866. }
    stellte eine bisher unbewiesene, aber auch nicht widerlegte Formel
    für die Verteilung von Primzahlen auf, die die Abschätzung 
    weiter verbessern würde.

\item Das Benfordsche Gesetz\footnote{%
        \url{http://de.wikipedia.org/wiki/Benfordsches_Gesetz}, \\
        \url{http://www.spiegel.de/wissenschaft/mensch/0,1518,632541,00.html},\\
        \url{http://arxiv.org/PS_cache/arxiv/pdf/0906/0906.2789v1.pdf}. }
    gilt nicht für Primzahlen.

    Nach dem Benfordschen Gesetz sind die Ziffern
    in den Zahlen bestimmter empirischer Datensätze (z.B. über Einwohnerzahlen
    von Städten, Geldbeträge in der Buchhaltung, Naturkonstanten) ungleichmäßig
    verteilt: Z.B. ist die Ziffer \verb#1# viel häufiger die erste Ziffer
    einer Zahl als jede andere.

    Welche Datensätze diesem Gesetz gehorchen ist noch nicht vollständig geklärt.
    Timo Eckhardt untersuchte in seiner Diplomarbeit 2008 ausführlich Eigenschaften
    von Primzahlen. Unter anderem wurden alle Primzahlen bis 7.052.046.499 mit
    verschiedenen Stellenwert-Basen dargestellt.

    Beim Vergleich der Basen 3 bis 10 ergab sich, dass die Abweichung von Benfords
    Gesetz bei der Basis 3 am geringsten ist. Für die Basis zehn besteht in etwa eine
    Gleichverteilung der ersten Ziffern. Bei der Untersuchung größerer Basen ergab sich,
    dass die Verteilung der ersten Ziffern von Basis zu Basis sehr starke Unterschiede
    aufweist.

@@

Primzahlzwillinge sind Primzahlen, die den Abstand 2 voneinander haben, 
zum Beispiel 5 und 7, oder 101 und 103,
oder $1.693.965 \cdot 2^{66.443} \pm 1$.

### 2009-11-23 script/en/chapters/ellipticcurves.tex
Script[de+en]: Small additions like research about Benford's law and primes.

Since we may define different operations on the same set, we distinguish them by giving them different names (e.g. $+$ addition or $\cdot$ multiplication).

@@

Fields may contain an infinite number of elements (e.g. the field of real numbers). They are called {\em infinite} fields. In contrast we call a field
{\em finite}, if it contains only a finite number of elements (e.g. ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$).
\index{Field!characteristic}
\paragraph{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. Then the characteristic of $K$ is said to be the order of $1$ with respect to the additive operation. This means that the characteristic of $K$ is the smallest positive integer $n$ such that
$$ \underbrace{1+1+\dots+1}_{\hbox{$n$ times}} =0 .
$$
If there is no such $n$, i.e. if $1+1+\dots+1\ne 0$ no matter how many $1$s we add, then we call $K$ a field
with characteristic $0$.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  thw "Elliptic Curve Cryptography (ECC) Tutorial"\\
  \url{http://www.williamstein.org/simuw06/notes/notes/node12.html}

@@

Sage notebooks are running after a logon within a browser%
\footnote{%
If you installed Sage on your own (Unix) server, you first have to enter at the
command line the command \verb#notebook()#.
},\footnote{%
The ECC notebook of Mrs. Massierer needs the KASH3 library: Therefore e.g. with Sage 4.1
the package ``kash3-2008-07-31.spkg'' has to be compiled.
}.

The notebook of Massierer\footnote{%
If you log on at another public server like
\url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/} or
\url{http://www.sagenb.org/}, and if there the according worksheets are installed
as "Published Worksheets", you can proceed as follows:\\
- Start the Sage notebook in a browser.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Click in the above right corner at ``Published'', then you will see the 8 worksheets.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sometimes it's necessary at the beginning to click in the left upper corner
  \verb#Action -> Evaluate all#.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  This tutorial uses the fonts from jsMath. They should be installed into your
  browser for a better layout.
  See \url{http://www.math.union.edu/~dpvc/jsMath/}
  and \url{http://pubpages.unh.edu/~jsh3/jsMath/}.\\
  After installing these fonts you can see the jsMath symbol at the lower border
  of your browser. If you click this symbol you can find the download page for the TIFF fonts.
  We are sorry, that this has to be done at every PC where his notebook runs in a browser.

@@

},\footnote{%
  Since 2008 this tutorial can be found at
    \url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}.\\
  To logon in Siegen you have to allow port 8000 and Cookies.\\
}
consists of 8 chapters (title page plus 7 chapters)
% "Titelseite" mit Inhaltsverzeichnis plus 7 Kapitel
and aims to let even beginners understand what elliptic curves are:

>>>>

Since we may define different operations on the same set, we distinguish them by giving them different names (e.g.\ $+$ addition or $\cdot$ multiplication).

@@

Fields may contain an infinite number of elements (e.g.\ the field of real numbers). They are called {\em infinite} fields. In contrast we call a field
{\em finite}, if it contains only a finite number of elements (e.g.\ ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$).
\index{Field!characteristic}
\paragraph{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. Then the characteristic of $K$ is said to be the order of $1$ with respect to the additive operation. This means that the characteristic of $K$ is the smallest positive integer $n$ such that
$$ \underbrace{1+1+\dots+1}_{\hbox{$n$ times}} =0 .
$$
If there is no such $n$, i.e. if $1+1+\dots+1\ne 0$ no matter how many $1$s we add, then we call $K$ a field
with characteristic $0$.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  the "Elliptic Curve Cryptography (ECC) Tutorial"\\
  \url{http://www.williamstein.org/simuw06/notes/notes/node12.html}

@@

Sage notebooks are running after a logon within a browser%
\footnote{%
If you installed Sage on your own (Unix) server, you first have to enter at the
command line the command \verb#notebook()#.
}${}^,$\footnote{%
The ECC notebook of Maike Massierer needs the KASH3 library: Therefore e.g.\ with Sage 4.2.1
the package ``kash3-2008-07-31.spkg'' has to be installed (command \verb#sage -i#).
}.

The ECC notebook\index{Elliptic curves!ECC notebook}\index{Massierer, Maike} of
Massierer\footnote{%
If you log on at another public server like
\url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/} or
\url{http://www.sagenb.org/}, and if there the according worksheets are installed
as "Published Worksheets", you can proceed as follows:\\
- Start the Sage notebook in a browser.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Click in the above right corner on ``Published'', then you will see the 8 worksheets.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sometimes it's necessary at the beginning to click in the left upper corner on
  \verb#Action -> Evaluate all#.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  This tutorial uses the fonts from jsMath. They should be installed into your
  browser for a better layout.
  See \url{http://www.math.union.edu/~dpvc/jsMath/}
  and \url{http://pubpages.unh.edu/~jsh3/jsMath/}.\\
  After installing these fonts you can see the jsMath symbol at the lower border
  of your browser. If you click this symbol you can find the download page for
  the TIFF fonts. We are sorry, that this installation has to be done at every
  PC where this notebook runs in a browser.

@@

}${}^,$\footnote{%
  Since 2008 this ECC notebook can be found at
    \url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}.\\
  To logon in Siegen you have to allow port 8000 and Cookies.\\
}
consists of 8 parts (title page plus 7 chapters)
% "Titelseite" mit Inhaltsverzeichnis plus 7 Kapitel
and aims to let even beginners understand what elliptic curves are:

### 2009-11-23 script/en/chapters/numbertheory.tex
Script[de+en]: Small additions like research about Benford's law and primes.

For more details on prime numbers (e.g. how ``Fermat's Little Theorem'' can
be used to test extremely large numbers to determine whether they are prime),
please refer to the article on \hyperlink{Kapitel_2}{prime numbers, 
chapter~\ref{Label_Kapitel_2}} in this script.

@@

In general mathematicians investigate ``Structures''\index{Structure}. They ask e.g. at $ a * x \equiv b $ mod $m$,
which values $x$ can take for given values of $a, ~b, ~m.$

@@

\item[{\bf 2.}] Select an arbitrary $e \in \{2, \cdots, n-1\}$ such
                that\footnote{%
                \label{foot:Selection-of-e}%
                It is recommended by cryptanalytic reasons, but not
                necessary to make RSA work, to select $e$ such that: \\
                $\max(p,q)  <  e <  J(n) - 1$.}: \\
                $e$ is relatively prime to $J(n) = (p-1)*(q-1)$. \\
                We can then ``throw away'' $p$ and $q$.\footnote{%
                The procedure also allows us to select $d$ freely and then 
                calculate $e$. However, this has practical disadvantages.
                We usually want to be able to encrypt messages ``quickly'',
                which is why we choose a public exponent $e$ such that 
                it has a short bit length compared to the modulus $n$ and 
                as few binary ones as possible (e.g. $2^{16} + 1$).
                So a fast exponentiation is possible when encrypting.                
                We want to select the publicly known $e$ to be an
                advantageous value that allows the exponential calculation
                to be performed quickly during encryption. The prime numbers
                $3, 17$ and $65537$ have proved to be particularly practical
                for this purpose.
                The most often used number is $65537 = 2^{16}+1$, or in
                binary: 
                $10\cdots 0\cdots 01$ (this number is prime and therefore
                relatively prime to many other numbers).
                }

@@

According attacks based on lattice reductions\index{Lattice reduction} 
are a real threat for (too) simple implementations of RSA.
Theses attacks can be structured into the following four categories:
\begin{itemize}
   \item Attacks against very small public keys e (e.g. $e = 3$). 
   \item Attacks against relatively small private exponents d(e.g. $d < N ^ {0.5}$).
   \item Factorization of the modulus N, if one of the factors p or q is {\em partly} known.
   \item Attacks requiring, that a {\em part} of the private key d is known.
\end{itemize}

@@

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
E.g. using CrypTool\index{CrypTool} via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorization of a Number}. \\
CrypTool can factorize in a reasonable time numbers no longer than 250 bit.
Numbers bigger than 1024 bits are currently not accepted by CrypTool.
}.
The next three prime factors were found by P.~Zimmerman\footnote{%
\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}}, 
T.~Grandlund\footnote{\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/}}
and R. Harley during the years 1999 and 2000 using the elliptic curve factorization method.

@@

The 3-page article in the DuD issue of June 2003 \cite{nt:Weis2003} contains
a very good explanation, how the attack using the Generalized Number Field
Sieve (GNFS) \index{General Number Field Sieve (GNFS)} works and which 
progress is made, to factorize numbers.
At GNFS we can distinguish 2 general steps: 
The sieve step (relation collecting) and the matrix reduction.
Besides the sieve step is highly parallelizable, it dominates the overall
calculation burden. Shamir and Tromer haven't built a TWIRL device yet,
but the estimated costs of 10 till 50 million Euro (in order to factorize
a 1024-bit number) is not prohibitive for secret agencies or big criminal
organizations, because the ``costs for a single espionage satellite is
estimated e.g. to be several billion USD''. The authors therefore
recommend, to get as soon as possible rid of today used sensible RSA, 
Diffie-Hellman or ElGamal keys up to 1024 bit and to use then keys of at
least 2048 bit length.
The planned TCPA/Palladium hardware \index{Palladium} will use 2048-bit
RSA keys!

@@

\vskip +20 pt
\subsection*{Number of digits to represent a positive integer}
For a positive integer $z$ the length of the $b$-adic representation can be determined
via the following steps. Starting from the inequality $b^{n+1} > z \ge b^n$ 
we have --- after applying the logarithm function on basis $b$\footnote{%
Applying the logarithm formula on base $b$ and $b'$ we have
$\log_b z = \log_{b'} z / \log_{b'} (b)$. 
It is therefore easy using e.g. logarithm tables for the base $b' = 10$ to compute
the logarithm of base $b = 2$.
}
:~$n+1 > log_b z \ge n$. Therefore we have $n = \lfloor log_b z \rfloor$\footnote{%
The function $\lfloor x \rfloor$ determines the next integer smaller than $x$
(in case $x \ge 0$ the digits after the decimal point are truncated). 
}
. We call $l_b(z)$ {\em the number of required digits to represent the number
$z$ on the base $b$}. We have $$l_b(z) := \lfloor log_b z \rfloor +1.$$

>>>>

For more details on prime numbers (e.g.\ how ``Fermat's Little Theorem'' can
be used to test extremely large numbers to determine whether they are prime),
please refer to the article on \hyperlink{Kapitel_2}{prime numbers, 
chapter~\ref{Label_Kapitel_2}} in this script.

@@

In general mathematicians investigate ``Structures''\index{Structure}. They ask e.g.\ at $ a * x \equiv b $ mod $m$,
which values $x$ can take for given values of $a, ~b, ~m.$

@@

\item[{\bf 2.}] Select an arbitrary $e \in \{2, \cdots, n-1\}$ such
                that\footnote{%
                \label{foot:Selection-of-e}%
                It is recommended by cryptanalytic reasons, but not
                necessary to make RSA work, to select $e$ such that: \\
                $\max(p,q)  <  e <  J(n) - 1$.}: \\
                $e$ is relatively prime to $J(n) = (p-1)*(q-1)$. \\
                We can then ``throw away'' $p$ and $q$.\footnote{%
                The procedure also allows us to select $d$ freely and then 
                calculate $e$. However, this has practical disadvantages.
                We usually want to be able to encrypt messages ``quickly'',
                which is why we choose a public exponent $e$ such that 
                it has a short bit length compared to the modulus $n$ and 
                as few binary ones as possible (e.g.\ $2^{16} + 1$).
                So a fast exponentiation is possible when encrypting.                
                We want to select the publicly known $e$ to be an
                advantageous value that allows the exponential calculation
                to be performed quickly during encryption. The prime numbers
                $3, 17$ and $65537$ have proved to be particularly practical
                for this purpose.
                The most often used number is $65537 = 2^{16}+1$, or in
                binary: 
                $10\cdots 0\cdots 01$ (this number is prime and therefore
                relatively prime to many other numbers).
                }

@@

According attacks based on lattice reductions\index{Lattice reduction} 
are a real threat for (too) simple implementations of RSA.
Theses attacks can be structured into the following four categories:
\begin{itemize}
   \item Attacks against very small public keys e (e.g.\ $e = 3$). 
   \item Attacks against relatively small private exponents d(e.g.\ $d < N ^ {0.5}$).
   \item Factorization of the modulus N, if one of the factors p or q is {\em partly} known.
   \item Attacks requiring, that a {\em part} of the private key d is known.
\end{itemize}

@@

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
E.g.\ using CrypTool\index{CrypTool} via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorization of a Number}. \\
CrypTool can factorize in a reasonable time numbers no longer than 250 bit.
Numbers bigger than 1024 bits are currently not accepted by CrypTool.
}.
The next three prime factors were found by P.~Zimmerman\footnote{%
\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}}, 
T.~Grandlund\footnote{\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/}}
and R. Harley during the years 1999 and 2000 using the elliptic curve factorization method.

@@

The 3-page article in the DuD issue of June 2003 \cite{nt:Weis2003} contains
a very good explanation, how the attack using the Generalized Number Field
Sieve (GNFS) \index{General Number Field Sieve (GNFS)} works and which 
progress is made, to factorize numbers.
At GNFS we can distinguish 2 general steps: 
The sieve step (relation collecting) and the matrix reduction.
Besides the sieve step is highly parallelizable, it dominates the overall
calculation burden. Shamir and Tromer haven't built a TWIRL device yet,
but the estimated costs of 10 till 50 million Euro (in order to factorize
a 1024-bit number) is not prohibitive for secret agencies or big criminal
organizations, because the ``costs for a single espionage satellite is
estimated e.g.\ to be several billion USD''. The authors therefore
recommend, to get as soon as possible rid of today used sensible RSA, 
Diffie-Hellman or ElGamal keys up to 1024 bit and to use then keys of at
least 2048 bit length.
The planned TCPA/Palladium hardware \index{Palladium} will use 2048-bit
RSA keys!

@@

\vskip +20 pt
\subsection*{Number of digits to represent a positive integer}
For a positive integer $z$ the length of the $b$-adic representation can be determined
via the following steps. Starting from the inequality $b^{n+1} > z \ge b^n$ 
we have --- after applying the logarithm function on basis $b$\footnote{%
Applying the logarithm formula on base $b$ and $b'$ we have
$\log_b z = \log_{b'} z / \log_{b'} (b)$. 
It is therefore easy using e.g.\ logarithm tables for the base $b' = 10$ to compute
the logarithm of base $b = 2$.
}
:~$n+1 > log_b z \ge n$. Therefore we have $n = \lfloor log_b z \rfloor$\footnote{%
The function $\lfloor x \rfloor$ determines the next integer smaller than $x$
(in case $x \ge 0$ the digits after the decimal point are truncated). 
}
. We call $l_b(z)$ {\em the number of required digits to represent the number
$z$ on the base $b$}. We have $$l_b(z) := \lfloor log_b z \rfloor +1.$$

### 2009-11-23 script/en/chapters/paper_and_pencil.tex
Script[de+en]: Small additions like research about Benford's law and primes.

\item {\bf Scytale}\footnote{%
   This method can directly be found in CrypTool\index{CrypTool} at the menu item
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Scytale / Rail Fence}.
   As this method is a special case of a simple columnar transposition, you also can
   simulate it in CrypTool\index{CrypTool} under the 
   menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
   Permutation}: For the Scytale within the dialog box only the first 
   permutation is used. If the wood has e.g. 4 angles use as key ``1,2,3,4''.
   This is equivalent to write the text horizontally in blocks of 4 letters 
   in a matrix and to read it out vertically . 
   Because the key is in an in ascending order, the Scytale is denoted as
   an identical permutation. And because writing and read-out is done only
   once it is a simple (and no double) permutation.}
   \cite{pp:Singh2001}\index{Scytale}%
   : 
   This method was probably used since 600 B.C. -- a description
   of how it operated is not known from before Plutarch (50-120 B.C.).\\
   A long strip of paper is wrapped around a wooden cylinder and then the 
   message is written along the length of this strip. The ciphertext is 
   produced by unwinding the strip.

@@

\item {\bf Baconian Cipher} \cite{pp:ACA2002}\index{Baconian Cipher}: 
   Assign a five-digit binary code to
   every letter and to 6 numbers or special characters (for example 00000 = A,
   00001 = B, etc.) and replace the plaintext characters with this binary code.
   Now use a second, unsuspicious message to hide the ciphertext inside 
   of it. This may happen by upper and lower case or italicized letters: 
   e.g. all letters of the unsuspicious message below  a binary ``1'' are 
   capitalised.

@@

\noindent To make the sample code\footnote{%
     Further examples with Sage can be found e.g. at:\\
     \href{http://www.sagemath.org/doc/constructions/linear_codes.html\#classical-ciphers}{\tt
           http://www.sagemath.org/doc/constructions/linear\_codes.html\#classical-ciphers}
                                          }
easier to understand, we used the structure and the naming conventions
shown in the graphics below:
\begin{itemize}
  \item Encryption consists of the two steps encoding and enciphering.
  \begin{itemize}
    \item Encoding adapts the letters in the given plaintext P to the case defined in the
          given alphabet, and all non-alphabet characters are filtered out.
    \item Enciphering creates the ciphertext C.
  \end{itemize}
  \item Decryption also consists of two steps: deciphering and decoding.
  \begin{itemize}
    \item Decoding is only necessary if the symbols in the alphabet are not ASCII characters.
  \end{itemize}
\end{itemize}

>>>>

\item {\bf Scytale}\footnote{%
   This method can directly be found in CrypTool\index{CrypTool} at the menu item
   {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} Scytale / Rail Fence}.
   As this method is a special case of a simple columnar transposition, you also can
   simulate it in CrypTool\index{CrypTool} under the 
   menu {\bf Crypt/Decrypt \textbackslash{} Symmetric (classic) \textbackslash{} 
   Permutation}: For the Scytale within the dialog box only the first 
   permutation is used. If the wood has e.g.\ 4 angles use as key ``1,2,3,4''.
   This is equivalent to write the text horizontally in blocks of 4 letters 
   in a matrix and to read it out vertically . 
   Because the key is in an in ascending order, the Scytale is denoted as
   an identical permutation. And because writing and read-out is done only
   once it is a simple (and no double) permutation.}
   \cite{pp:Singh2001}\index{Scytale}%
   : 
   This method was probably used since 600 B.C. -- a description
   of how it operated is not known from before Plutarch (50-120 B.C.).\\
   A long strip of paper is wrapped around a wooden cylinder and then the 
   message is written along the length of this strip. The ciphertext is 
   produced by unwinding the strip.

@@

\item {\bf Baconian Cipher} \cite{pp:ACA2002}\index{Baconian Cipher}: 
   Assign a five-digit binary code to
   every letter and to 6 numbers or special characters (for example 00000 = A,
   00001 = B, etc.) and replace the plaintext characters with this binary code.
   Now use a second, unsuspicious message to hide the ciphertext inside 
   of it. This may happen by upper and lower case or italicized letters: 
   e.g.\ all letters of the unsuspicious message below  a binary ``1'' are 
   capitalised.

@@

\noindent To make the sample code\footnote{%
     Further examples with Sage can be found e.g.\ at:\\
     \href{http://www.sagemath.org/doc/constructions/linear_codes.html\#classical-ciphers}{\tt
           http://www.sagemath.org/doc/constructions/linear\_codes.html\#classical-ciphers}
                                          }
easier to understand, we used the structure and the naming conventions
shown in the graphics below:
\begin{itemize}
  \item Encryption consists of the two steps encoding and enciphering.
  \begin{itemize}
    \item Encoding adapts the letters in the given plaintext P to the case defined in the
          given alphabet, and all non-alphabet characters are filtered out.
    \item Enciphering creates the ciphertext C.
  \end{itemize}
  \item Decryption also consists of two steps: deciphering and decoding.
  \begin{itemize}
    \item Decoding is only necessary if the symbols in the alphabet are not ASCII characters.
  \end{itemize}
\end{itemize}
